Cover letter for the IJRR submission \\ ``An efficient acyclic contact planner for multiped robots''

### 2017-01-23 body/efort.tex
Review section 5
complete paper
no symbol redefinition

% !TEX root =  ../main.tex
\subsection{Manipulability-based heuristics for contact selection}
This Section proposes heuristics to select a contact that optimizes desired capabilities.
%~ These heuristics allow to locally optimize a contact location with respect to a particular task.
For instance, one can be interested in configurations that allow to efficiently exert a force in the global direction of motion,
a high velocity in a given direction, or to stay away from singular configurations.
We derive three such heuristics from the work by \cite{Yoshikawa1984}, recalled here. %, that we present first.

@@

\begin{figure}[!tbp]
  \centering
	\begin{overpic}[width=1\linewidth]{figures/EFORT/ellipsoid}
		\put (8.5,1.2) {\small{Velocity ellipsoid}}
		\put (62.5,1.2) {\small{Force ellipsoid} \tiny{(scale 0.5)}}
	\end{overpic}
  \caption{Examples of velocity and force ellipsoids for a manipulator composed of 2 dofs and 2 segments.
Only the horizontal and vertical speeds are shown (not the rotation speeds).}
 %~ since it would require being able to draw in four dimensions.}
		   \label{sec:efort_ellipsoid}
\end{figure}

@@

With EFORT, we define the efficiency of a configuration as the ability of a limb to exert a force in a given direction.
We thus consider the force ellipsoid as a basis for our heuristic.
In a given direction $\mathbf{m}$, the length of the ellipsoid is given by the force-transmission ratio \citep{1087795}:
\begin{equation*}
f_\mathsf{T}(\mathbf{q}, \mathbf{m}) = [\mathbf{m}^{T}(\mathbf{J}\mathbf{J}	^{T})\mathbf{m}]^{-\frac{1}{2}}
\end{equation*}

In our problem, to compare candidate configurations, we include the quality of the contact surface, and choose $\mathbf{m}$ as the direction
opposite to the local motion (given by the difference between two consecutive root positions):

\begin{equation}
h_{\textrm{\it EFORT}}(\mathbf{q}, \mathbf{m}) = [\mathbf{m}^{T}(\mathbf{J}\mathbf{J}^T)\mathbf{m}]^{-\frac{1}{2}} ( \mu \mathbf{n}^T \mathbf{m})
\end{equation}
where $\mu$ and $\mathbf{n}$ are respectively the friction coefficient and the normal vector of the contact surface.

If the ability to generate large velocities at the effector is considered, we define a new heuristic $h_{vel}$ with a similar reasoning on the velocity ellipsoid:
\begin{equation}
h_{vel}(\mathbf{q}, \mathbf{m}) = [\mathbf{m}^{T}(\mathbf{J}\mathbf{J}^T)^{-1}\mathbf{m}]^{-\frac{1}{2}} ( \mu \mathbf{n}^T \mathbf{m})
\end{equation}

>>>>

\section{Heuristics and criteria for contact selection}
\label{sec:heuristics}
% !TEX root =  ../main.tex
\subsection{Manipulability-based heuristics for contact selection}
This Section proposes heuristics to select a contact that optimizes desired capabilities.
%~ These heuristics allow to locally optimize a contact location with respect to a particular task.
For instance, one can be interested in configurations that allow to efficiently exert a force in the global direction of motion,
a high velocity in a given direction, or to stay away from singular configurations.
We derive three such heuristics from the work by \cite{Yoshikawa1984}, recalled here. %, that we present first.

@@

\begin{figure}[!tbp]
  \centering
	\begin{overpic}[width=1\linewidth]{figures/EFORT/ellipsoid}
		\put (8.5,1.2) {\small{Velocity ellipsoid}}
		\put (62.5,1.2) {\small{Force ellipsoid} \tiny{(scale 0.5)}}
	\end{overpic}
  \caption{Examples of velocity and force ellipsoids for a manipulator composed of 2 DoFs and 2 segments.
Only the horizontal and vertical speeds are shown (not the rotation speeds).}
 %~ since it would require being able to draw in four dimensions.}
		   \label{sec:efort_ellipsoid}
\end{figure}

@@

With EFORT, we define the efficiency of a configuration as the ability of a limb to exert a force in a given direction.
We thus consider the force ellipsoid as a basis for our heuristic.
In a given direction $\uprho$, the length of the ellipsoid is given by the force-transmission ratio \citep{1087795}:
\begin{equation*}
f_\mathsf{T}(\mathbf{q}, \uprho) = [\uprho^{T}(\mathbf{J}\mathbf{J}	^{T})\uprho]^{-\frac{1}{2}}
\end{equation*}

In our problem, to compare candidate configurations, we include the quality of the contact surface, and choose $\uprho$ as the direction
opposite to the local motion (given by the difference between two consecutive root positions):

\begin{equation}
h_{\textrm{\it EFORT}}(\mathbf{q}, \uprho) = [\uprho^{T}(\mathbf{J}\mathbf{J}^T)\uprho]^{-\frac{1}{2}} ( \mu \mathbf{n}^T \uprho)
\end{equation}
where $\mu$ and $\mathbf{n}$ are respectively the friction coefficient and the normal vector of the contact surface.

If the ability to generate large velocities at the effector is considered, we define a new heuristic $h_{vel}$ with a similar reasoning on the velocity ellipsoid:
\begin{equation}
h_{vel}(\mathbf{q}, \uprho) = [\uprho^{T}(\mathbf{J}\mathbf{J}^T)^{-1}\uprho]^{-\frac{1}{2}} ( \mu \mathbf{n}^T \uprho)
\end{equation}

### 2017-01-23 body/equilibrium.tex
complete paper
no symbol redefinition

% !TEX root =  ../main.tex
\section{Heuristics and criteria for contact selection}
\label{sec:heuristics}
\subsection{A criterion for robust static equilibrium}
The planner is designed so that any generated contact configuration is in static equilibrium.
%~ Equilibrium is of course critical in legged locomotion. For this reason 
We are interested in a robust 
criterion, that ensures that the robot remains in equilibrium in a real-world application, regardless of perception and control uncertainties.

@@

\subsubsection{Conditions for static equilibrium:}
We first define the variables of the problem, for $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
%~ \item $\mathbf{L}  \in \mathbb{R}^3 $ is the angular momentum at the COM;
\item $m \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact point $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{t}_{i1},\mathbf{t}_{i2}$ form a local Cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{t}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

According to Coulomb's law, the nonslipping condition is verified if all the contact forces lie in the friction cone defined by the surface.
As classically done, we linearize the friction cone in a conservative fashion with a pyramid, described by four generating rays of unit length. We choose for instance:
\begin{equation*}
\mathbf{V}_{i} = \mat{\mathbf{n}_{i} + \mu \mathbf{t}_{i1} & \mathbf{n}_{i} -\mu \mathbf{t}_{i1} & \mathbf{n}_{i} + \mu \mathbf{t}_{i2} & \mathbf{n}_{i} - \mu \mathbf{t}_{i2}}^T
\end{equation*}

>>>>

% !TEX root =  ../main.tex

\subsection{A criterion for robust static equilibrium}
The planner is designed so that any generated contact configuration is in static equilibrium.
%~ Equilibrium is of course critical in legged locomotion. For this reason 
We are interested in a robust 
criterion, that ensures that the robot remains in equilibrium in a real-world application, regardless of perception and control uncertainties.

@@

\subsubsection{Conditions for static equilibrium:}
We first define the variables of the problem, for $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
%~ \item $\mathbf{L}  \in \mathbb{R}^3 $ is the angular momentum at the COM;
\item $m \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact point $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{\boldsymbol\upgamma}_{i1},\mathbf{\boldsymbol\upgamma}_{i2}$ form a local Cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{\boldsymbol\upgamma}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

According to Coulomb's law, the nonslipping condition is verified if all the contact forces lie in the friction cone defined by the surface.
As classically done, we linearize the friction cone in a conservative fashion with a pyramid, described by four generating rays of unit length. We choose for instance:
\begin{equation*}
\mathbf{V}_{i} = \mat{\mathbf{n}_{i} + \mu \mathbf{\boldsymbol\upgamma}_{i1} & \mathbf{n}_{i} -\mu \mathbf{\boldsymbol\upgamma}_{i1} & \mathbf{n}_{i} + \mu \mathbf{\boldsymbol\upgamma}_{i2} & \mathbf{n}_{i} - \mu \mathbf{\boldsymbol\upgamma}_{i2}}^T
\end{equation*}
